---
title: "GnRH_TRH_CCK. Analyse PAS between drug and time measurements"
author: "Yaromir Kobikov <kobikov.yaromir@gmail.com>"
date: "04/13/2025"
output:
 word_document: default
 html_document:
  toc: true
  toc_float:
   collapsed: false
editor_options:
 chunk_output_type: inline
---

```{r setup, include=FALSE}
source("../BayesianPValue.R")
```

# PAS (Magnitude of Postural asymmetry)


# Read Data
Read from Excel file.

```{r input}

rds_folder_name <- "../rds/task_1/pas/"

if (!dir.exists(rds_folder_name)) {
  dir.create(rds_folder_name, recursive = TRUE)
}

file_name <- "../data/GnRH_TRH_CCK_Stat_25 04 10.xlsx"
sheet <- "formatted"
my.seed <- 103592321 # reproducible Bayessian fit
myname0 <- "PAS"

data <-
  read_excel(file_name, sheet, .name_repair = "universal", na = c("", "NA", "NULL")) %>%
  select(
    Exp.ID, 
    BW,
    Trt,
    Dose,
    matches("^Hand\\.(1|2|3|4)\\.\\d+$"),
    matches("^MP\\.(1|2|3|4)\\.\\d+$")
  ) %>%
  rename(
    ExpID = `Exp.ID`
  ) %>%
  mutate(
    across(where(is.character), ~ str_remove_all(., "\\s")),
    Dose = str_to_lower(str_trim(Dose)),
    dose.relative_to_n = case_when(
      # If the dose is "0", assign numeric zero.
      Dose == "1" ~ 1,
      # If dose ends with 'n': remove any spaces plus the trailing "n" and convert to numeric.
      str_detect(Dose, "n$") ~ suppressWarnings(as.numeric(str_remove(Dose, "\\s*n$"))),
      # If dose ends with 'p': remove any spaces plus the trailing "p" and convert to numeric, then multiply by 1e-3 (1p = 0.001 n)
      str_detect(Dose, "p$") ~ suppressWarnings(as.numeric(str_remove(Dose, "\\s*p$"))) * 1e-3,
      # Otherwise, assign NA
      TRUE ~ NA_real_
    ),
    # Log10 transformation: log10(0) is undefined, so return 0 for dose.relative_to_n <= 0
    Dose.Log = if_else(dose.relative_to_n > 0, log10(dose.relative_to_n), NA_real_)
  ) %>%
  relocate(Dose.Log, .after = Dose) %>% 
  pivot_longer(
    cols = c(Hand.1.1:Hand.4.3, MP.1.1:MP.4.3),
    names_to = c("Measurement method", "Time"),
    names_pattern = "(Hand|MP).(.)",
    values_to = "PA"
  ) %>%
  filter(Time %in% c("1", "4")) %>% 
  mutate(RatID = factor(ExpID),
         Time = factor(Time, c("1", "4"), c("0min", "180min")),
         Drug6 = if_else(Trt == 'TRH', paste0(Trt, '.', Dose), Trt),
         Drug6 = factor(Drug6, c("CCK8", "GnRH", "Saline", "TRH.0.1n", "TRH.10n", "TRH.1n")),
         Drug = factor(Trt, c("Saline", "TRH", "GnRH", "CCK8")),
         MPA = abs(PA)) %>%
  drop_na("PA") %>%
  ungroup %>% droplevels(.)

data_hand <-
  data %>%
  filter(`Measurement method` == "Hand") %>%
  droplevels(.)

data_machine <-
  data %>%
  filter(`Measurement method` == "MP") %>%
  droplevels(.)

```

The group sizes are (distinct rats only, same rats with different stimulation locations are counted as duplicates)

```{r input2, echo=FALSE}
data %>%
  arrange(Drug, Time) %>%
  group_by(Drug, Time, `Measurement method`) %>%
  summarise(rats = length(unique(RatID)), .groups = "drop") %>%
  pivot_wider(names_from=c(Time), values_from=rats) %>%
  flextable %>%
  autofit

```

The drug (+ by dose) sizes are (distinct rats only, same rats with different stimulation locations are counted as duplicates)

```{r input2_group_by_dose, echo=FALSE}

data %>%
  arrange(Drug, Time) %>%
  group_by(Drug, Time, `Measurement method`, Dose) %>%
  summarise(rats = length(unique(RatID)), .groups = "drop") %>%
  pivot_wider(names_from=c(Time), values_from=rats) %>%
  flextable %>%
  autofit

```

# (Hand) Analyse Drug dose:

## PAS: Student Model

Prior predictive check

```{r hand_check_prior_dose}

myname <- paste0(myname0, "_Hand")

std.prior <- brm(
  data = data_hand,
  family = student,
  PA ~ 0 + Time:Drug + Drug*Dose.Log + (1|RatID),
  prior = c(prior(student_t(3, 0, 3), class = "sigma"),
            prior(normal(0, 2), class = "b")),
  seed = my.seed,
  chains = my.cores,
  cores = my.cores,
  iter = 4e4,
  sample_prior = "only",
  file = paste0(rds_folder_name, "check_prior_check_dose.rds"),
  file_refit = "on_change"
)

cond_eff <- conditional_effects(std.prior,
                                "Drug",
                                method="fitted",
                                conditions=make_conditions(data_hand, vars = c("Time", "Dose.Log")))

plt[["Drug*Dose.Log PAS prior vs data (hand)"]] <- plot(cond_eff,
                                       points = TRUE,
                                       point_args = list(width = 0.1, shape="o", col="red"),
                                       plot=FALSE)[[1]] +
                                  geom_hline(yintercept=0, linetype="dashed", color="black")

set.seed(2)
plt[["Drug*Dose.Log PAS prior pp_check (hand)"]] <- pp_check(std.prior, ndraws = 11)

print(cond_eff[[1]])
print(plt[["Drug*Dose.Log PAS prior vs data (hand)"]])
print(plt[["Drug*Dose.Log PAS prior pp_check (hand)"]])

```

Model fit and QC:

```{r hand_model_summary_dose, echo=FALSE, fig.width = 10}

std.fit <-  update(std.prior,
                   sample_prior = "yes",
                   control = list(max_treedepth = 15),
                   seed = my.seed,
                   cores = my.cores,
                   file = paste0(rds_folder_name, "model_summary_dose.rds"),
                   file_refit = "on_change")
summary(std.fit)

```

## PAS: MCMC conversion diagnostics

Posterior predictive check

```{r hand_pas_MCMC_ppcheck_dose}

set.seed(2) # reproducible pp_check

plt[["Drug*Dose.Log PAS pp_check (hand)"]] <- pp_check(std.fit, ndraws = 21)
print(plt[["Drug*Dose.Log PAS pp_check (hand)"]])

```

Auto correlations

```{r hand_pas_MCMC_ac_dose}

plt[["Drug*Dose.Log PAS ac (hand)"]] <- stan_ac(std.fit$fit)
print(plt[["Drug*Dose.Log PAS ac (hand)"]])

```

```{r trends_dose}
dose_trends <- emtrends(std.fit, specs = ~ Drug, var = "Dose.Log")
print(dose_trends)

dose_trends_contrasts <- pairs(dose_trends)
print(dose_trends_contrasts)

plt[["Dose_Log_Trends_PAS (hand)"]] <- dose_trends %>% 
  as.data.frame() %>%
  ggplot(aes(y = fct_rev(Drug), x = Dose.Log.trend, fill = Drug)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = lower.HPD, xmax = upper.HPD), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  labs(
    title = myname,
    x = "Estimated Dose.Log Trend (Slope)",
    y = "Drug"
  ) +
  theme_minimal()

print(plt[["Dose_Log_Trends_PAS (hand)"]])
```

verify explicitly with model comparison for TRH:

```{r verify_trh_1}
# fit_TRH_dose <- brm(
#   PA ~ Time + Dose.Log + (1|RatID),
#   data = subset(data_hand, Drug == "TRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# fit_TRH_no_dose <- brm(
#   PA ~ Time + (1|RatID),
#   data = subset(data_hand, Drug == "TRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# loo_compare(loo(fit_TRH_dose), loo(fit_TRH_no_dose))
```

                elpd_diff se_diff
fit_TRH_dose     0.0       0.0   
fit_TRH_no_dose -0.6       0.9   

Expected Log Predictive Density difference
Standard error of that difference
abs(elpd_diff) > 2 × se_diff

The model including dose (fit_TRH_dose) is considered slightly better (elpd_diff=0), but the difference (-0.6) is very small relative to the uncertainty (se_diff = 0.9).
This indicates essentially no meaningful improvement from including Dose.Log.

verify explicitly with model comparison for GnRH:

```{r verify_gnrh}
# fit_GnRH_dose <- brm(
#   PA ~ Time + Dose.Log + (1|RatID),
#   data = subset(data_hand, Drug == "GnRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# fit_GnRH_no_dose <- brm(
#   PA ~ Time + (1|RatID),
#   data = subset(data_hand, Drug == "GnRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# loo_compare(loo(fit_GnRH_dose), loo(fit_GnRH_no_dose))
```

                 elpd_diff se_diff
fit_GnRH_no_dose  0.0       0.0   
fit_GnRH_dose    -0.3       1.6  

The simpler model (fit_GnRH_no_dose) without Dose.Log is slightly favored.
However, the difference is very small (-0.3) and clearly not statistically meaningful given the large uncertainty (se_diff = 1.6).
This indicates no evidence that including Dose.Log improves your model.

verify explicitly with model comparison for TRH test 2:

```{r verify_trh2}
# fit_TRH_dose2 <- brm(
#   PA ~ 0 + Time + Dose.Log + (1|RatID),
#   data = subset(data_hand, Drug == "TRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# fit_TRH_no_dose2 <- brm(
#   PA ~ 0 + Time + (1|RatID),
#   data = subset(data_hand, Drug == "TRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# loo_compare(loo(fit_TRH_dose2), loo(fit_TRH_no_dose2))
```

                 elpd_diff se_diff
fit_TRH_dose2     0.0       0.0   
fit_TRH_no_dose2 -0.7       1.0   

verify explicitly with model comparison for GnRH test 2:

```{r verify_gnrh2}
# fit_GnRH_dose2 <- brm(
#   PA ~ 0 + Time + Dose.Log + (1|RatID),
#   data = subset(data_hand, Drug == "GnRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# fit_GnRH_no_dose2 <- brm(
#   PA ~ 0 + Time + (1|RatID),
#   data = subset(data_hand, Drug == "GnRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# loo_compare(loo(fit_GnRH_dose2), loo(fit_GnRH_no_dose2))
```

                  elpd_diff se_diff
fit_GnRH_no_dose2  0.0       0.0   
fit_GnRH_dose2    -0.3       1.6   

verify explicitly with model comparison for Drug:

```{r verify_trh_2}
# fit_dose <- brm(
#   PA ~ 0 + Time:Drug + Drug*Dose.Log + (1|RatID),
#   data = data_hand,
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# fit_no_dose <- brm(
#   PA ~ 0 + Time:Drug + (1|RatID),
#   data = data_hand,
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# loo_compare(loo(fit_dose), loo(fit_no_dose))
```

            elpd_diff se_diff
fit_no_dose  0.0       0.0   
fit_dose    -1.5       2.2   

## PAS: Estimated model marginals Time | Drug
Estimated model medians +- 95% HPDCI (HPDCI = highest posterior density continuous interval) as a table:

```{r hand_pas_emm_dose}

emm <- emmeans(
  std.fit, 
  specs = ~ Time | Drug,
  type = "response"
)
emm_show(emm)

```

Estimated model means as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):

```{r hand_pas_emm_plot_dose, echo=FALSE, fig.width = 8}

plt[["Drug*Dose.Log Time|Drug PAS emm (hand)"]] <- emm %>% emm_inorder %>%
  ggplot(aes(y = fct_rev(interaction(Time, Drug, sep=", ", drop=TRUE)), x = .value, fill = Time)) %>%
  my.stat_eyeh() +
  geom_vline(xintercept=0, linetype="dashed", color="black") +
  # geom_vline(xintercept=threshold, linetype="dashed", color="black") +
  # geom_vline(xintercept=-threshold, linetype="dashed", color="black") +
  labs(title=myname, y="", x="PAS (mm) Time|Drug + Drug*Dose.Log")

print(plt[["Drug*Dose.Log Time|Drug PAS emm (hand)"]])

```

Estimated model means as a median +- 95% HPDCI ("x" - mean values of measurements for each rat)

```{r points_hand_pas_emm_plot_dose, echo=FALSE, fig.width=8}

plt[["Drug*Dose.Log Time|Drug PAS emm (hand) with Data Points"]] <-
  plt[["Drug*Dose.Log Time|Drug PAS emm (hand)"]] +
    geom_point(
      pch = 4,
      position = position_jitterdodge(dodge.width = 0.6, jitter.width = 0.6, seed = 2),
      data = {
      data_hand %>%
        group_by(RatID, Time, Drug) %>% 
        summarise(
          .value = mean(PA),
          .groups = "drop"
        )
    },
      aes(x=.value, y = as.numeric(fct_rev(interaction(Time, Drug, sep=", ", drop=TRUE)))+0.2))

print(plt[["Drug*Dose.Log Time|Drug PAS emm (hand) with Data Points"]])

```
## PAS: Contrasts by Drug

```{r hand_pas_contrasts_dose_drug}

emmc <- rbind(contrast(emm, simple = "Time", method = "revpairwise", reverse = TRUE))
emm_show(emmc)

```

Contrasts as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):

```{r hand_pas_plot_dose_drug, echo=FALSE, fig.width = 8}

plt[["Drug*Dose.Log Time|Drug PAS by Drug (hand)"]] <- emmc %>% emm_inorder %>%
  ggplot(aes( y = fct_rev(interaction(Drug, contrast, sep=", ", drop=TRUE)), x=.value, fill=contrast)) %>%
  my.stat_eyeh() +
  geom_vline(xintercept=0, linetype="dashed", color="black") +
  labs(title=myname, y="", x="PAS in mm, contrast: Time")

print(plt[["Drug*Dose.Log Time|Drug PAS by Drug (hand)"]])


```

## PAS: Contrasts by Time

```{r hand_pas_contrasts_dose_time}

emmc <- rbind(contrast(emm, simple = "Drug", method = "revpairwise", reverse = TRUE))
emm_show(emmc)

```

Contrasts as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):

```{r hand_pas_plot_dose_time, echo=FALSE, fig.width = 8}

plt[["Drug*Dose.Log Time|Drug PAS by Time (hand)"]] <- emmc %>% emm_inorder %>%
  ggplot(aes( y = fct_rev(interaction(Time, contrast, sep=", ", drop=TRUE)), x=.value, fill=contrast)) %>%
  my.stat_eyeh() +
  geom_vline(xintercept=0, linetype="dashed", color="black") +
  labs(title=myname, y="", x="PAS in mm, contrast: Drug")

print(plt[["Drug*Dose.Log Time|Drug PAS by Time (hand)"]])

```

The Drug6 (without dose) sizes are (distinct rats only, same rats with different stimulation locations are counted as duplicates)

```{r input_drug6, echo=FALSE}
data %>%
  arrange(Drug6, Time) %>%
  group_by(Drug6, Time, `Measurement method`) %>%
  summarise(rats = length(unique(RatID)), .groups = "drop") %>%
  pivot_wider(names_from=c(Time), values_from=rats) %>%
  flextable %>%
  autofit

```

# (Hand) Analyse Drug|Time (without Dose):
## PAS: Student Model
Prior predictive check

```{r hand_check_prior}

myname <- paste0(myname0, "_Hand")

std.prior <- brm(
  data = data_hand,
  family = student,
  PA ~ 0 + Time:Drug6 + (1|RatID),
  prior = c(prior(student_t(3, 0, 3), class = "sigma"),
            prior(normal(0, 2), class = "b")),
  seed = my.seed,
  chains = my.cores,
  cores = my.cores,
  iter = 4e4,
  sample_prior = "only",
  file = paste0(rds_folder_name, "check_prior_check.rds"),
  file_refit = "on_change"
)

cond_eff <- conditional_effects(std.prior,
                                "Drug6",
                                method="fitted",
                                conditions=make_conditions(data_hand, vars = c("Time")))

plt[["PAS prior vs data (hand)"]] <- plot(cond_eff,
                                       points = TRUE,
                                       point_args = list(width = 0.1, shape="o", col="red"),
                                       plot=FALSE)[[1]] +
                                  geom_hline(yintercept=0, linetype="dashed", color="black")

set.seed(2)
plt[["PAS prior pp_check (hand)"]] <- pp_check(std.prior, ndraws = 11)

print(cond_eff[[1]])
print(plt[["PAS prior vs data (hand)"]])
print(plt[["PAS prior pp_check (hand)"]])

```

Model fit and QC:
```{r hand_model_summary, echo=FALSE, fig.width = 10}

std.fit <-  update(std.prior,
                   sample_prior = "yes",
                   control = list(max_treedepth = 15),
                   seed = my.seed,
                   cores = my.cores,
                   file = paste0(rds_folder_name, "model_summary.rds"),
                   file_refit = "on_change")
summary(std.fit)

```

## PAS: MCMC conversion diagnostics
Posterior predictive check
```{r hand_pas_MCMC_ppcheck}

set.seed(2) # reproducible pp_check

plt[["PAS pp_check (hand)"]] <- pp_check(std.fit, ndraws = 21)
print(plt[["PAS pp_check (hand)"]])

```

Auto correlations
```{r hand_pas_MCMC_ac}

plt[["PAS ac (hand)"]] <- stan_ac(std.fit$fit)
print(plt[["PAS ac (hand)"]])

```

## PAS: Estimated model marginals Time | Drug
Estimated model medians +- 95% HPDCI (HPDCI = highest posterior density continuous interval) as a table:
```{r hand_pas_emm}

emm <- emmeans(
  std.fit, 
  specs = ~ Time | Drug6,
  type = "response"
)
emm_show(emm)

```

Estimated model means as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):
```{r hand_pas_emm_plot, echo=FALSE, fig.width = 8}

plt[["Time|Drug6 PAS emm (hand)"]] <- emm %>% emm_inorder %>%
  ggplot(aes(y = fct_rev(interaction(Time, Drug6, sep=", ", drop=TRUE)), x = .value, fill = Time)) %>%
  my.stat_eyeh() +
  geom_vline(xintercept=0, linetype="dashed", color="black") +
  # geom_vline(xintercept=threshold, linetype="dashed", color="black") +
  # geom_vline(xintercept=-threshold, linetype="dashed", color="black") +
  labs(title=myname, y="", x="PAS (mm) Time|Drug6")

print(plt[["Time|Drug6 PAS emm (hand)"]])

```

Estimated model means as a median +- 95% HPDCI ("x" - mean values of measurements for each rat)
```{r points_hand_pas_emm_plot, echo=FALSE, fig.width=8}

plt[["Time|Drug6 PAS emm (hand) with Data Points"]] <-
  plt[["Time|Drug6 PAS emm (hand)"]] +
    geom_point(
      pch = 4,
      position = position_jitterdodge(dodge.width = 0.6, jitter.width = 0.6, seed = 2),
      data = {
      data_hand %>%
        group_by(RatID, Time, Drug6) %>% 
        summarise(
          .value = mean(PA),
          .groups = "drop"
        )
    },
      aes(x=.value, y = as.numeric(fct_rev(interaction(Time, Drug6, sep=", ", drop=TRUE)))+0.2))

print(plt[["Time|Drug6 PAS emm (hand) with Data Points"]])

```

## PAS: Contrasts by Drug
```{r hand_pas_contrasts_drug}

emmc <- rbind(contrast(emm, simple = "Time", method = "revpairwise", reverse = TRUE))
emm_show(emmc)

```

Contrasts as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):
```{r hand_pas_plot_drug, echo=FALSE, fig.width = 8}

plt[["Time|Drug6 PAS by Drug (hand)"]] <- emmc %>% emm_inorder %>%
  ggplot(aes( y = fct_rev(interaction(Drug6, contrast, sep=", ", drop=TRUE)), x=.value, fill=contrast)) %>%
  my.stat_eyeh() +
  geom_vline(xintercept=0, linetype="dashed", color="black") +
  labs(title=myname, y="", x="PAS in mm, contrast: Time")

print(plt[["Time|Drug6 PAS by Drug (hand)"]])

```

## PAS: Contrasts by Time

```{r hand_pas_contrasts_time}

emmc <- rbind(contrast(emm, simple = "Drug6", method = "revpairwise", reverse = TRUE))
emm_show(emmc)

```

Contrasts as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):

```{r hand_pas_plot_time, echo=FALSE, fig.width = 8}

plt[["Time|Drug6 PAS by Time (hand)"]] <- emmc %>% emm_inorder %>%
  ggplot(aes( y = fct_rev(interaction(Time, contrast, sep=", ", drop=TRUE)), x=.value, fill=contrast)) %>%
  my.stat_eyeh() +
  geom_vline(xintercept=0, linetype="dashed", color="black") +
  labs(title=myname, y="", x="PAS in mm, contrast: Drug6")

print(plt[["Time|Drug6 PAS by Time (hand)"]])

```

















==========================================================================================================
# (Machine) Comparison between measurements (Time) and groups (Group)

## PAS: Student Model
Prior predictive check
```{r machine_check_prior_dose}

myname <- paste0(myname0, "_Machine")

std.prior <- brm(
  data = data_machine,
  family = student,
  PA ~ 0 + Time:Drug + Drug*Dose.Log + (1|RatID),
  prior = c(prior(student_t(3, 0, 3), class = "sigma"),
            prior(normal(0, 2), class = "b")),
  seed = my.seed,
  chains = my.cores,
  cores = my.cores,
  iter = 4e4,
  sample_prior = "only",
  file = paste0(rds_folder_name, "check_prior_check_dose_machine.rds"),
  file_refit = "on_change"
)

cond_eff <- conditional_effects(std.prior,
                                "Drug",
                                method="fitted",
                                conditions=make_conditions(data_hand, vars = c("Time", "Dose.Log")))

plt[["Drug*Dose.Log PAS prior vs data (machine)"]] <- plot(cond_eff,
                                       points = TRUE,
                                       point_args = list(width = 0.1, shape="o", col="red"),
                                       plot=FALSE)[[1]] +
                                  geom_hline(yintercept=0, linetype="dashed", color="black")

set.seed(2)
plt[["Drug*Dose.Log PAS prior pp_check (machine)"]] <- pp_check(std.prior, ndraws = 11)

print(cond_eff[[1]])
print(plt[["Drug*Dose.Log PAS prior vs data (machine)"]])
print(plt[["Drug*Dose.Log PAS prior pp_check (machine)"]])

```

Model fit and QC:

```{r machine_model_summary_dose, echo=FALSE, fig.width = 10}

std.fit <-  update(std.prior,
                   sample_prior = "yes",
                   control = list(max_treedepth = 15),
                   seed = my.seed,
                   cores = my.cores,
                   file = paste0(rds_folder_name, "model_summary_dose_machine.rds"),
                   file_refit = "on_change")
summary(std.fit)

```

## PAS: MCMC conversion diagnostics
Posterior predictive check
```{r machine_pas_MCMC_ppcheck_dose}

set.seed(2) # reproducible pp_check

plt[["Drug*Dose.Log PAS pp_check (machine)"]] <- pp_check(std.fit, ndraws = 21)
print(plt[["Drug*Dose.Log PAS pp_check (machine)"]])

```

Auto correlations

```{r machine_pas_MCMC_ac_dose}

plt[["Drug*Dose.Log PAS ac (machine)"]] <- stan_ac(std.fit$fit)
print(plt[["Drug*Dose.Log PAS ac (machine)"]])

```

```{r machine_trends_dose}
dose_trends <- emtrends(std.fit, specs = ~ Drug, var = "Dose.Log")
print(dose_trends)

dose_trends_contrasts <- pairs(dose_trends)
print(dose_trends_contrasts)

plt[["Dose_Log_Trends_PAS (machine)"]] <- dose_trends %>% 
  as.data.frame() %>%
  ggplot(aes(y = fct_rev(Drug), x = Dose.Log.trend, fill = Drug)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = lower.HPD, xmax = upper.HPD), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  labs(
    title = myname,
    x = "Estimated Dose.Log Trend (Slope)",
    y = "Drug"
  ) +
  theme_minimal()

print(plt[["Dose_Log_Trends_PAS (machine)"]])
```

verify explicitly with model comparison for TRH:

```{r machine_verify_trh}
# fit_TRH_dose <- brm(
#   PA ~ Time + Dose.Log + (1|RatID),
#   data = subset(data_machine, Drug == "TRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# fit_TRH_no_dose <- brm(
#   PA ~ Time + (1|RatID),
#   data = subset(data_machine, Drug == "TRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# loo_compare(loo(fit_TRH_dose), loo(fit_TRH_no_dose))
```
                elpd_diff se_diff
fit_TRH_dose     0.0       0.0   
fit_TRH_no_dose -0.3       0.6   

Expected Log Predictive Density difference
Standard error of that difference
abs(elpd_diff) > 2 × se_diff

The model including dose (fit_TRH_dose) is considered slightly better (elpd_diff=0), but the difference (-0.6) is very small relative to the uncertainty (se_diff = 0.9).
This indicates essentially no meaningful improvement from including Dose.Log.

verify explicitly with model comparison for GnRH:

```{r machine_verify_gnrh}
# fit_GnRH_dose <- brm(
#   PA ~ Time + Dose.Log + (1|RatID),
#   data = subset(data_machine, Drug == "GnRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# fit_GnRH_no_dose <- brm(
#   PA ~ Time + (1|RatID),
#   data = subset(data_machine, Drug == "GnRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# loo_compare(loo(fit_GnRH_dose), loo(fit_GnRH_no_dose))
```

                 elpd_diff se_diff
fit_GnRH_dose     0.0       0.0   
fit_GnRH_no_dose -0.4       0.9  

The simpler model (fit_GnRH_no_dose) without Dose.Log is slightly favored.
However, the difference is very small (-0.3) and clearly not statistically meaningful given the large uncertainty (se_diff = 1.6).
This indicates no evidence that including Dose.Log improves your model.

verify explicitly with model comparison for TRH test 2:

```{r machine_verify_trh2}
# fit_TRH_dose2 <- brm(
#   PA ~ 0 + Time + Dose.Log + (1|RatID),
#   data = subset(data_machine, Drug == "TRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# fit_TRH_no_dose2 <- brm(
#   PA ~ 0 + Time + (1|RatID),
#   data = subset(data_machine, Drug == "TRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# loo_compare(loo(fit_TRH_dose2), loo(fit_TRH_no_dose2))
```

                 elpd_diff se_diff
fit_GnRH_dose     0.0       0.0   
fit_GnRH_no_dose -0.4       0.9  

verify explicitly with model comparison for GnRH test 2:

```{r machine_verify_gnrh2}
# fit_GnRH_dose2 <- brm(
#   PA ~ 0 + Time + Dose.Log + (1|RatID),
#   data = subset(data_machine, Drug == "GnRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# fit_GnRH_no_dose2 <- brm(
#   PA ~ 0 + Time + (1|RatID),
#   data = subset(data_machine, Drug == "GnRH"),
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# loo_compare(loo(fit_GnRH_dose2), loo(fit_GnRH_no_dose2))
```

                  elpd_diff se_diff
fit_GnRH_dose2     0.0       0.0   
fit_GnRH_no_dose2 -0.3       1.0   

verify explicitly with model comparison for Drug:

```{r machine_verify_trh_3}
# fit_dose <- brm(
#   PA ~ 0 + Time:Drug + Drug*Dose.Log + (1|RatID),
#   data = data_machine,
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# fit_no_dose <- brm(
#   PA ~ 0 + Time:Drug + (1|RatID),
#   data = data_machine,
#   family = student(),
#   prior = c(prior(student_t(3, 0, 3), class = "sigma"),
#             prior(normal(0, 2), class = "b")),
#   seed = my.seed,
#   chains = my.cores,
#   cores = my.cores,
#   iter = 4e4,
# )
# 
# loo_compare(loo(fit_dose), loo(fit_no_dose))
```

            elpd_diff se_diff
fit_no_dose  0.0       0.0   
fit_dose    -1.9       1.9  


## PAS: Estimated model marginals Time | Drug
Estimated model medians +- 95% HPDCI (HPDCI = highest posterior density continuous interval) as a table:

```{r machine_pas_emm_dose}

emm <- emmeans(
  std.fit, 
  specs = ~ Time | Drug,
  type = "response"
)
emm_show(emm)

```

Estimated model means as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):

```{r machine_pas_emm_plot_dose, echo=FALSE, fig.width = 8}

plt[["Drug*Dose.Log Time|Drug PAS emm (machine)"]] <- emm %>% emm_inorder %>%
  ggplot(aes(y = fct_rev(interaction(Time, Drug, sep=", ", drop=TRUE)), x = .value, fill = Time)) %>%
  my.stat_eyeh() +
  geom_vline(xintercept=0, linetype="dashed", color="black") +
  # geom_vline(xintercept=threshold, linetype="dashed", color="black") +
  # geom_vline(xintercept=-threshold, linetype="dashed", color="black") +
  labs(title=myname, y="", x="PAS (mm) Time|Drug + Drug*Dose.Log (machine)")

print(plt[["Drug*Dose.Log Time|Drug PAS emm (machine)"]])

```

Estimated model means as a median +- 95% HPDCI ("x" - mean values of measurements for each rat)

```{r points_machine_pas_emm_plot_dose, echo=FALSE, fig.width=8}

plt[["Drug*Dose.Log Time|Drug PAS emm (machine) with Data Points"]] <-
  plt[["Drug*Dose.Log Time|Drug PAS emm (machine)"]] +
    geom_point(
      pch = 4,
      position = position_jitterdodge(dodge.width = 0.6, jitter.width = 0.6, seed = 2),
      data = {
      data_machine %>%
        group_by(RatID, Time, Drug) %>% 
        summarise(
          .value = mean(PA),
          .groups = "drop"
        )
    },
      aes(x=.value, y = as.numeric(fct_rev(interaction(Time, Drug, sep=", ", drop=TRUE)))+0.2))

print(plt[["Drug*Dose.Log Time|Drug PAS emm (machine) with Data Points"]])

```

## PAS: Contrasts by Drug

```{r machine_pas_contrasts_dose_drug}

emmc <- rbind(contrast(emm, simple = "Time", method = "revpairwise", reverse = TRUE))
emm_show(emmc)

```

Contrasts as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):

```{r machine_pas_plot_dose_drug, echo=FALSE, fig.width = 8}

plt[["Drug*Dose.Log Time|Drug PAS by Drug (machine)"]] <- emmc %>% emm_inorder %>%
  ggplot(aes( y = fct_rev(interaction(Drug, contrast, sep=", ", drop=TRUE)), x=.value, fill=contrast)) %>%
  my.stat_eyeh() +
  geom_vline(xintercept=0, linetype="dashed", color="black") +
  labs(title=myname, y="", x="PAS in mm, contrast: Time (machine)")

print(plt[["Drug*Dose.Log Time|Drug PAS by Drug (machine)"]])


```

## PAS: Contrasts by Time

```{r machine_pas_contrasts_dose_time}

emmc <- rbind(contrast(emm, simple = "Drug", method = "revpairwise", reverse = TRUE))
emm_show(emmc)

```

Contrasts as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):

```{r machine_pas_plot_dose_time, echo=FALSE, fig.width = 8}

plt[["Drug*Dose.Log Time|Drug PAS by Time (machine)"]] <- emmc %>% emm_inorder %>%
  ggplot(aes( y = fct_rev(interaction(Time, contrast, sep=", ", drop=TRUE)), x=.value, fill=contrast)) %>%
  my.stat_eyeh() +
  geom_vline(xintercept=0, linetype="dashed", color="black") +
  labs(title=myname, y="", x="PAS in mm, contrast: Drug (machine)")

print(plt[["Drug*Dose.Log Time|Drug PAS by Time (machine)"]])

```

The Drug6 (without dose) sizes are (distinct rats only, same rats with different stimulation locations are counted as duplicates)

```{r machine_input, echo=FALSE}
data %>%
  arrange(Drug6, Time) %>%
  group_by(Drug6, Time, `Measurement method`) %>%
  summarise(rats = length(unique(RatID)), .groups = "drop") %>%
  pivot_wider(names_from=c(Time), values_from=rats) %>%
  flextable %>%
  autofit

```

# (machine) Analyse Drug|Time (without Dose):
## PAS: Student Model
Prior predictive check

```{r machine_check_prior}

myname <- paste0(myname0, "_Machine")

std.prior <- brm(
  data = data_machine,
  family = student,
  PA ~ 0 + Time:Drug6 + (1|RatID),
  prior = c(prior(student_t(3, 0, 3), class = "sigma"),
            prior(normal(0, 2), class = "b")),
  seed = my.seed,
  chains = my.cores,
  cores = my.cores,
  iter = 4e4,
  sample_prior = "only",
  file = paste0(rds_folder_name, "check_prior_check_machine.rds"),
  file_refit = "on_change"
)

cond_eff <- conditional_effects(std.prior,
                                "Drug6",
                                method="fitted",
                                conditions=make_conditions(data_hand, vars = c("Time")))

plt[["PAS prior vs data (machine)"]] <- plot(cond_eff,
                                       points = TRUE,
                                       point_args = list(width = 0.1, shape="o", col="red"),
                                       plot=FALSE)[[1]] +
                                  geom_hline(yintercept=0, linetype="dashed", color="black")

set.seed(2)
plt[["PAS prior pp_check (machine)"]] <- pp_check(std.prior, ndraws = 11)

print(cond_eff[[1]])
print(plt[["PAS prior vs data (machine)"]])
print(plt[["PAS prior pp_check (machine)"]])

```

Model fit and QC:
```{r machine_model_summary, echo=FALSE, fig.width = 10}

std.fit <-  update(std.prior,
                   sample_prior = "yes",
                   control = list(max_treedepth = 15),
                   seed = my.seed,
                   cores = my.cores,
                   file = paste0(rds_folder_name, "model_summary_machine.rds"),
                   file_refit = "on_change")
summary(std.fit)

```

## PAS: MCMC conversion diagnostics
Posterior predictive check
```{r machine_pas_MCMC_ppcheck}

set.seed(2) # reproducible pp_check

plt[["PAS pp_check (machine)"]] <- pp_check(std.fit, ndraws = 21)
print(plt[["PAS pp_check (machine)"]])

```

Auto correlations
```{r machine_pas_MCMC_ac}

plt[["PAS ac (machine)"]] <- stan_ac(std.fit$fit)
print(plt[["PAS ac (machine)"]])

```

## PAS: Estimated model marginals Time | Drug
Estimated model medians +- 95% HPDCI (HPDCI = highest posterior density continuous interval) as a table:
```{r machine_pas_emm}

emm <- emmeans(
  std.fit, 
  specs = ~ Time | Drug6,
  type = "response"
)
emm_show(emm)

```

Estimated model means as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):
```{r machine_pas_emm_plot, echo=FALSE, fig.width = 8}

plt[["Time|Drug6 PAS emm (machine)"]] <- emm %>% emm_inorder %>%
  ggplot(aes(y = fct_rev(interaction(Time, Drug6, sep=", ", drop=TRUE)), x = .value, fill = Time)) %>%
  my.stat_eyeh() +
  geom_vline(xintercept=0, linetype="dashed", color="black") +
  # geom_vline(xintercept=threshold, linetype="dashed", color="black") +
  # geom_vline(xintercept=-threshold, linetype="dashed", color="black") +
  labs(title=myname, y="", x="PAS (mm) Time|Drug6 (machine)")

print(plt[["Time|Drug6 PAS emm (machine)"]])

```

Estimated model means as a median +- 95% HPDCI ("x" - mean values of measurements for each rat)
```{r points_machine_pas_emm_plot, echo=FALSE, fig.width=8}

plt[["Time|Drug6 PAS emm (machine) with Data Points"]] <-
  plt[["Time|Drug6 PAS emm (machine)"]] +
    geom_point(
      pch = 4,
      position = position_jitterdodge(dodge.width = 0.6, jitter.width = 0.6, seed = 2),
      data = {
      data_machine %>%
        group_by(RatID, Time, Drug6) %>% 
        summarise(
          .value = mean(PA),
          .groups = "drop"
        )
    },
      aes(x=.value, y = as.numeric(fct_rev(interaction(Time, Drug6, sep=", ", drop=TRUE)))+0.2))

print(plt[["Time|Drug6 PAS emm (machine) with Data Points"]])

```

## PAS: Contrasts by Drug
```{r machine_pas_contrasts_drug}

emmc <- rbind(contrast(emm, simple = "Time", method = "revpairwise", reverse = TRUE))
emm_show(emmc)

```

Contrasts as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):
```{r machine_pas_plot_drug, echo=FALSE, fig.width = 8}

plt[["Time|Drug6 PAS by Drug (machine)"]] <- emmc %>% emm_inorder %>%
  ggplot(aes( y = fct_rev(interaction(Drug6, contrast, sep=", ", drop=TRUE)), x=.value, fill=contrast)) %>%
  my.stat_eyeh() +
  geom_vline(xintercept=0, linetype="dashed", color="black") +
  labs(title=myname, y="", x="PAS in mm, contrast: Time (machine)")

print(plt[["Time|Drug6 PAS by Drug (machine)"]])

```

## PAS: Contrasts by Time

```{r machine_pas_contrasts_time}

emmc <- rbind(contrast(emm, simple = "Drug6", method = "revpairwise", reverse = TRUE))
emm_show(emmc)

```

Contrasts as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):

```{r machine_pas_plot_time, echo=FALSE, fig.width = 8}

plt[["Time|Drug6 PAS by Time (machine)"]] <- emmc %>% emm_inorder %>%
  ggplot(aes( y = fct_rev(interaction(Time, contrast, sep=", ", drop=TRUE)), x=.value, fill=contrast)) %>%
  my.stat_eyeh() +
  geom_vline(xintercept=0, linetype="dashed", color="black") +
  labs(title=myname, y="", x="PAS in mm, contrast: Drug6 (machine)")

print(plt[["Time|Drug6 PAS by Time (machine)"]])

```







# Editable PowerPoint plots generated...

```{r pptx, message=TRUE, warning=TRUE, include=FALSE, results='hide'}

myname <- paste0(myname0, "_PAS_DrugTime_Hand_Machine")

doc <- read_pptx()

for(nm in names(plt)) {
  doc <- doc %>%
    add_slide(layout = "Title and Content", master = "Office Theme") %>%
    ph_with(value = rvg::dml(ggobj = plt[[nm]]),
          location = ph_location_type(type = "body"
                                      #, width=S_width, height=S_height
                                      ),
          bg = "transparent" ) %>%
    ph_with(value = nm,
          location = ph_location_type(type = "title") )
}
doc %>% print(target = paste0(myname,".pptx"))

```



